---
title: "MaRvelous MaRkdown; A seRies"
author: "EDNA MWENDA"
date: "9/24/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
R Markdown is a file format for making dynamic documents with R. An R Markdown document is written in markdown (an easy-to-write plain text format) and contains chunks of embedded R code.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```


## Let's get started!!

R Markdown supports dozens of static and dynamic output formats including HTML, PDF, MS Word, Beamer, HTML5 slides, Tufte-style handouts, books, dashboards, shiny applications, scientific articles, websites, and more. 


You can use a single R Markdown file to both

   - save and execute code
  - generate high quality reports that can be shared with an audience
    
##  How it works


When you run `render` , R Markdown feeds the .Rmd file to knitr, which executes all of the code chunks and creates a new markdown (.md) document which includes the code and its output.

The markdown file generated by knitr is then processed by pandoc which is responsible for creating the finished format.

This may sound complicated, but R Markdown makes it extremely simple by encapsulating all of the above processing into a single render function.


## Installation
 I assume you have already installed R (https://www.r-project.org) (R Core Team 2020) and the RStudio IDE (https://www.rstudio.com). RStudio is not required but recommended, because it makes it easier for an average user to work with R Markdown. If you do not have RStudio IDE installed, you will have to install Pandoc (http://pandoc.org), otherwise there is no need to install Pandoc separately because RStudio has bundled it. Next you can install the `rmarkdown` package in R:
```{r}
# Install from CRAN
#install.packages('rmarkdown')

# Or if you want to test the development version,
# install from GitHub
#if (!requireNamespace("devtools"))
  #install.packages('devtools')
#devtools::install_github('rstudio/rmarkdown')
```
 
If you want to generate PDF output, you will need to install LaTeX. For R Markdown users who have not installed LaTeX before, I recommend that you install TinyTeX (https://yihui.name/tinytex/):


```{r}
#install.packages('tinytex')
#tinytex::install_tinytex()  # install TinyTeX
```



##  TinyTex/ PDF format



TinyTeX is a lightweight, portable, cross-platform, and easy-to-maintain LaTeX distribution. 

The R companion package tinytex (Xie 2020e) can help you automatically install missing LaTeX packages when compiling LaTeX or R Markdown documents to PDF, and also ensures a LaTeX document is compiled for the correct number of times to resolve all cross-references.


## Applications of R Markdown in Real life


 -  Airbnb uses R Markdown to document all their analyses in R, so they can combine code and data visualizations in a single report.  Eventually all reports are carefully peer-reviewed and published to a company knowledge repository, so that anyone in the company can easily find analyses relevant to their team.
 
 -  R Markdown is used at eelloo to design and generate research reports.
 -  Used to build general-purpose websites (including blogs and personal websites) based on R Markdown.
 
 
 
## Not just for R!
 
 
 - Python
 
 -CSS, Javascript
 
 -Bash, Stan
 
 - SQL
 
 A total of 52 langauge engines from `knitr`, wooooow!!!
 
##  Inline Code

Code results can be inserted directly into the text of a .Rmd file by enclosing the code with `r `. 
Using `r` `package ` makes it easy to update the report to refer to another function.

##  Flexdashboards:
Easy interactive dashboards for R

First you install the `flexdashboard` package from CRAN as follows:
```{r}
#install.packages("flexdashboard")
library(flexdashboard, suppressMessages())
```

To author a flexdashboard you create an R Markdown document with the flexdashboard:: flex_dashboard output format as show in next(flexdashboard) document.

** Layout
Single column (Fill)

Dashboards are divided into columns, and rows, with output components delineated
using level 3 markdown headers(###)

Key is to play around with the `vertical_layout` function on YAML.

 For multiple columns you introduce a level 2 markdown header(-----) for each column. 
 
 ** Row Orientation

You can also choose to orient dashboards row-wise rather than column-wise by specifying the `orientation::rows` option.


##  Pandoc Markdown
Pandoc understands an extended and slightly revised version of John Gruber’s Markdown syntax. Markdown is designed to be easy to write, and, even more importantly, easy to read:

    A Markdown-formatted document should be publishable as-is, as plain text, without
    looking like it’s been marked up with tags or formatting instructions. – John Gruber
    
    
This principle has guided pandoc’s decisions in finding syntax for tables, footnotes, and other extensions.

There is, however, one respect in which pandoc’s aims are different from the original aims of Markdown. Whereas Markdown was originally designed with HTML generation in mind, pandoc is designed for multiple output formats. Thus, while pandoc allows the embedding of raw HTML, it discourages it, and provides other, non-HTMLish ways of representing important document elements like definition lists, tables, mathematics, and footnotes.
